---
title: "Simulating Diff in diff data using the R package simDAG"
format: html
editor: visual
---

## Sim DAG

```{r}
library("simDAG")
dag <- empty_dag() +
  node(c("A", "B"), type="rnorm", mean=0, sd=1) +
  node("C", type="gaussian", formula=~-2 + A*0.3 + B*-2,
       error=1)
```

```{r}
library("igraph")
library("ggplot2")
library("ggforce")
plot(dag, layout="as_tree", node_size=0.15, node_fill="grey",
     node_text_fontface="italic")
```

```{r}
summary(dag)
```

```{r}
library(tidyverse)
set.seed(123)
dt <- sim_from_dag(dag, n_sim = 100)

dt$Bb <- if_else(dt$B < 0, "red", "blue")
```

```{r}
ggplot(dt, aes(x = A, y = C, col = Bb)) + geom_point() + scale_colour_identity()
```

# Introducing multiple time points

```{r}
dag <- empty_dag() +
  node_td("Y", type="time_to_event", prob_fun=0.01,
          event_duration=Inf)

```

```{r}
sim <- sim_discrete_time(dag, n_sim=1000, max_t=80)
```

What is the probability that the event has not occured at t = 80 ?

This is (0.99)\^ 80

```{r}
0.99^80
```

So we expect on average 45% persons with event FALSE, and 55% persons with event TRUE.

```{r}
mean(sim$data$Y_event)
```

# Time varyi

```{r}
prob_myoc <- function(data, P_0, RR_A) {
  ifelse(data$A_event, P_0*RR_A, P_0)
}

dag <- empty_dag() +
  node_td("A", type="time_to_event", prob_fun=0.01,
          event_duration=20, immunity_duration=150) +
  node_td("Y", type="time_to_event", prob_fun=prob_myoc,
          parents=c("A_event"), P_0=0.005, RR_A=3.24)
```

```{r}
plot(dag)
```

```{r}
sim <- sim_discrete_time(dag, n_sim=10000, max_t=365*2)
```

```{r}
set.seed(123)
dat_ss <- sim2data(sim, to="start_stop", target_event="Y",
                keep_only_first=FALSE, overlap=TRUE)  # If TRUE, all information after the first event per person will be discarded. Useful when target_event should be treated as a terminal variable.
```

```{r}
dat_ss %>% filter(.id ==1)
```

So what we have here? A base probability of getting vaccinated

```{r}
# > head(dat) %>% filter(.id == 1)
#      .id start  stop      A      Y
#    <int> <int> <num> <lgcl> <lgcl>
# 1:     1     1   183  FALSE   TRUE
# 2:     1   183   225  FALSE   TRUE
# 3:     1   225   230  FALSE   TRUE
# 4:     1   230   264  FALSE   TRUE
# 5:     1   264   534  FALSE  FALSE
# 6:     1   534   554   TRUE  FALSE
```

```{r}
set.seed(123)
dat <- sim2data(sim, to="long") 
```

```{r}
dat %>% filter(.id ==1 & .time == 54) 
dat %>% filter(.id ==1 & .time == 293)

```

Great it matches.

```{r}
ggplot(dat %>% filter(.id ==1), aes(x = .time, y = A)) +geom_point()
```

So we expect that after A occurs (a person is risk elevated due to vaccination the first 20 days), the prob of the event is increased three fold.

So we can check this by subsetting the data on A==1, and checking the prob of Y in that period, compared to prob of Y otherwise.

There is also the immunity though, but that is immunity to vaccination, not to getting myoc.

```{r}
res <- dat %>%
  group_by(A) %>%
  summarise(P_Y = mean(Y))


RR_A_est <- res$P_Y[2]/res$P_Y[1]
RR_A_est
```

```{r}
res
```

```{r}
mean(dat$Y)
```

```{r}
res_t <- dat %>%
  group_by(.time) %>%
  summarise(P_A = mean(A),
            P_Y = mean(Y))


```

```{r}
t =0;for( i in 1:20 ){ t = t + (0.01 - i*0.0001)}
```

```{r}
ggplot(res_t %>% filter(.time < 20), aes(x = .time, y = P_A)) +
  geom_point() +
  geom_point(aes (y = P_Y), col = "blue") +
  geom_hline(aes(yintercept = mean(dat$Y)), col = "red") +
  geom_abline(aes(slope = 0.01, intercept = 0)) +
  geom_hline(aes(yintercept = t), col = "purple") #+
  #geom_vline(aes(xintercept = 20), col = "red") +
  #geom_vline(aes(xintercept = 150), col = "blue")
```

So what happens here? We have the base prob of getting vaccinated at 0.01. Because we have a duration of 20, we have a buildup to around 0.01 x 20. However at day 20, only 80 not vaccinated indiv are left. so the buildup slows down a bit

0.1 + 0.099 + 0.098 + 0.097 + 0.096 + 0.095

# What if we want data at multiple points in time?

# Actual DID datasets

Card and Krueger.

Medicaid expansion.
