---
title: "Simulate DiD data"
---

```{r}
library(tidyverse)
```

This paper simulate diff in dif data where patients at clinics get treated at the clinic level.

```{r}
source("../staggered-adoption-simstudy/fun-gen-params-2.R")
source("../staggered-adoption-simstudy/fun-gendata-3.R")
source("../staggered-adoption-simstudy/fun-methods.R")
```

We evaluated the methods on various simulation scenarios in which we varied the total number
of clusters, sample size, and the effect of exposure on the outcome. 

* The total sample sizes in the scenarios were **500**, 1000, 2000, and 5000. 

* The number of clusters also changed within each setting and was equal to **30**, 50, and 100. 

* The number of individuals per cluster was equally distributed. 

The effect of exposure on the outcome was modeled under three scenarios:
* constant (no heterogeneity), 
* lagged (temporal heterogeneity), 
* and heterogeneity by both group and time. 

The total number of time points was set to t = 5, and we enforced that each cluster start the intervention between periods 3 and 5, i.e. Gg = 0 for g = 1, 2. 

Hence, there were no never-treated units and two observed periods with no intervention effect.

We generated five covariates in total, three at the individual level and two at the cluster level.
Among the three individual covariates, two were continuous and one binary. 

These were meant to resemble our motivating example, in which the individual-level covariates are age, a frailty
score, and sex. 

The cluster-level covariates consisted of one categorical covariate with 3 levels and one binary covariate. 

Similarly, these covariates were meant to mimic clinic characteristics such as type of practice, and whether the clinic is located in an urban or rural area.

    # --- -- mu.t: vector of length equal to time.periods, time effects under no treatment
    # --- -- rho.t: vector of length equal to time.periods, time effects under treatment 

```{r}
n_participants <- 500
```


```{r}
# [1] "id"        "t"         "eps"       "nu"        "Y0"        "Y1"        "Y"         "Yg1"       "Yg2"       "Yg3"       "Yg4"       "Yg5"      
# [13] "treat"     "Gt"        "C"         "prtrt1"    "prtrt2"    "prtrt3"    "prtrt4"    "prtrt5"    "X1"        "X2"        "X3"        "X4"       
# [25] "V1"        "V2"        "V3"        "V4"        "V5"        "V6"        "I_1"       "I_2"       "I_3"       "I_4"       "I_5"       "I_6"      
# [37] "I_7"       "I_8"       "I_9"       "I_10"      "I_11"      "I_12"      "I_13"      "I_14"      "I_15"      "I_16"      "I_17"      "I_18"     
# [49] "I_19"      "I_20"      "I_21"      "I_22"      "I_23"      "I_24"      "I_25"      "I_26"      "I_27"      "I_28"      "I_29"      "I_30"     
# [61] "Gt_f"      "Gt_t"      "Gt_t_f"    "t_f"       "dist"      "label_dis" "label"     "dist_f"    "label_f
```


# Scenario 0: zero treatment effect, 30 clusters

In this scenario, we have 30 clusters (clinics), with total 500 participants. 
We have random effects by cluster. This is the psi.

We have 2 periods of pre-treatment. Cluster level covariates.

There is NO treatment effect.

participants cannot switch clusters.

```{r}
set.seed(123)

scen_params <- gen_params(scen = 0, np = n_participants)
res <- gen.data.cluster(scen_params)
pat_ids <- sample(1:n_participants, 10)
```

```{r}
ggplot(res$long_dat %>% filter(id %in% pat_ids), aes(x = t, y = Y, col = factor(C), group = id)) +
  
  geom_point() +
  geom_line() +
  geom_line(aes(y = treat*10-30, group = id), col = "blue")
```
If we would fit a fixed effect model we expect zero treatment effect.

```{r}
res$long_dat %>%
  group_by(treat, t) %>%
  summarise(avg_t = mean(Y1-Y0),
            n_obs = n())

```



# Scenario 1: constant treatment effect, single cluster

A single cluster. Constant treatment effect. All participants get treated t the same time.

```{r}
set.seed(123)

scen_params <- gen_params(scen = 1, np = n_participants)
res <- gen.data.cluster(scen_params)

pat_ids <- sample(1:n_participants, 10)
```

```{r}
ggplot(res$long_dat %>% filter(id %in% pat_ids), aes(x = t, y = Y, col = factor(C), group = id)) +
  
  geom_point() +
  geom_line() +
  geom_line(aes(y = treat*10, group = id), col = "blue")
```
```{r}
res$long_dat %>%
  group_by(treat, C) %>%
  summarise(avg_t = mean(Y1-Y0),
            n_obs = n())

```
```{r}
lm(Y ~ factor(t), data = res$long_dat)
```

# Scenario 2: time lagged treatment effect, single cluster

A single cluster. This particular cluster happens to get treated at T = 5 (seed 123).
What does time lag mean? 

```{r}
set.seed(145)

scen_params <- gen_params(scen = 2, np = n_participants)
res <- gen.data.cluster(scen_params)
pat_ids <- sample(1:n_participants, 10)
```

```{r}
ggplot(res$long_dat %>% filter(id %in% pat_ids), aes(x = t, y = Y, col = factor(C), group = id)) +
  
  geom_point() +
  geom_line() +
  geom_line(aes(y = treat*10, group = id), col = "blue")
```
```{r}
res$long_dat %>%
  group_by(treat, C) %>%
  summarise(avg_t = mean(Y1-Y0),
            n_obs = n())

```
So 2/3 of 2.

# Scenario 3 group-time heterogeneity, single cluster

There is only 1 group. effect increases in time.

```{r}
set.seed(145)

scen_params <- gen_params(scen = 3, np = n_participants)
res <- gen.data.cluster(scen_params)
pat_ids <- sample(1:n_participants, 10)
```

```{r}
ggplot(res$long_dat %>% filter(id %in% pat_ids), aes(x = t, y = Y, col = factor(C), group = id)) +
  
  geom_point() +
  geom_line() +
  geom_line(aes(y = treat*10, group = id), col = "blue")
```
```{r}
res$long_dat %>%
  group_by(treat, C, t) %>%
  summarise(avg_t = mean(Y1-Y0),
            n_obs = n())

```

# Scenario 4: only group heterogeneity, single cluster

With a single cluster, group heterogeneity is meaninglesS? For this cluster, treatment effect is 1.

```{r}
set.seed(145)

scen_params <- gen_params(scen = 4, np = n_participants)
res <- gen.data.cluster(scen_params)
pat_ids <- sample(1:n_participants, 10)
```

```{r}
ggplot(res$long_dat %>% filter(id %in% pat_ids), aes(x = t, y = Y, col = factor(C), group = id)) +
  
  geom_point() +
  geom_line() +
  geom_line(aes(y = treat*10, group = id), col = "blue")
```

```{r}
res$long_dat %>%
  group_by(treat, C, t) %>%
  summarise(avg_t = mean(Y1-Y0),
            n_obs = n())

```

# Scenario 5: constant treatment effect, ten clusters

```{r}
set.seed(123)

scen_params <- gen_params(scen = 5, np = n_participants)
res <- gen.data.cluster(scen_params)
pat_ids <- sample(1:n_participants, 10)
```

```{r}
ggplot(res$long_dat %>% filter(id %in% pat_ids), aes(x = t, y = Y, col = factor(C), group = id)) +
  
  geom_point() +
  geom_point(aes(y=Y0)) +
  geom_point(aes(y=Y1, shape=factor(treat))) +
  geom_line() +
  geom_line(aes(y = treat*10-25, group = id), col = "blue")
```
lets calculate the constant treatment effect from Y1 Y0.

```{r}
res$long_dat %>%
  group_by(treat, C) %>%
  summarise(avg_t = mean(Y1-Y0),
            n_obs = n())

```

